---
title: "Coursera Statistical Inference Project"
author: "Josh Starkey"
date: "January 31, 2016"
output: pdf_document
---

###Overview
This is a Coursera project for the Data Science Specialization: Statisical Inference Course.  The project is described https://www.coursera.org/learn/statistical-inference/peer/3k8j5/statistical-inference-course-project.  This project has two parts.  The first part investigates the exponential distribution in R and compares it with the Central Limit Theorem.  The second part performs some basic analysis on the ToothGrowth data provided by R.  This report is uploaded as a pdf, but the R Markdown is located https://github.com/star9475/StatInferenceProject/blob/master/StatInfProject.Rmd.

First, let's load libraries:
```{r, warning=FALSE, message=FALSE}
library(ggplot2)
```

###Part 1

####Simulations
Let's run a simulation of an exponential distribution with lambda = .2.

The expected mean of the distribution is 1/lambda.
```{r}
expmean <- 1 / .2
```
```{r, echo=FALSE}
expmean
```
The expected variance of the distribution is also 1/lambda.
```{r}
expvar <- ((1 / .2)^2)
```
```{r, echo=FALSE}
expvar
```
The number of simulations for the experiment is 1000.<br>
The number of samples is 40.
```{r}
nosim <- 1000
nosample <- 40
```
Let's perform 1000 simulations of the exponential distribution, and take 40 samples of each simulation.  We'll set the seed for reproducability.  
```{r, cache = TRUE, warning=FALSE,message=FALSE}
set.seed(1623);
meandata <- data.frame(
     x = c(apply(matrix(rexp(nosim * nosample, .2), nosim), 1, mean)),
     n = factor(rep(c(nosample), c(nosample))) 
     )
```

Let's perform the same simulation: 1000 simulations of the exponential distribution.  This time we'll calculate the variance of the samples.  We'll set the seed for reproducability.  
```{r, cache = TRUE, warning=FALSE,message=FALSE}
set.seed(1623);
vardata <- data.frame(
     x = c(apply(matrix(rexp(nosim * nosample, .2), nosim), 1, var)),
     n = factor(rep(c(nosample), c(nosample))) 
     )
```

####Sample Mean versus Theoretical Mean

We plot the distribution of sample means.  The red line represents the expected mean <code>expmean = 5</code>, and the green line represents the sample mean<code>sampmean = mean(meandata[,1])=5.011907</code>.  They are very close.<br>
```{r,  fig.align="left", cache = TRUE, warning=FALSE,message=FALSE, echo=FALSE}
g <- ggplot(data = meandata, aes(x = x) ) +
     geom_histogram(aes(y = ..density..), fill = ("blue"), col=("darkgreen"), alpha=.4) +
     labs(title ="Average Means of 40 Samples", x="means") +
     geom_vline(xintercept=expmean, color="red") +
     geom_vline(xintercept=mean(meandata[,1]), color="green")
g
```

####Sample Variance versus Theoretical Variance

We plot the distribution of sample variances.  The red line represents the expected variance <code>expvar = 25</code>, and the green line represents the sample variance<code>sampvar = mean(vardata[,1])=24.91826</code>.  They are very close.<br>
```{r, fig.align="left", cache = TRUE, warning=FALSE,message=FALSE, echo=FALSE}
g <- ggplot(data = vardata, aes(x = x) ) +
     geom_histogram(aes(y = ..density..), fill = ("blue"), col=("darkgreen"), alpha=.4) +
     labs(title ="Average Vars of 40 Samples", x="Vars") +
     geom_vline(xintercept=expvar, color="red") + 
     geom_vline(xintercept=mean(vardata[,1]), color="green")
g
```

####Distribution
The Central Limit Theorm states that averages are approximately normal with distributions centered at the population mean, which we have shown in the first figure "Average Means of 40 samples".  

###Part 2

####Load the ToothGrowth data and perform some basic exploratory data analyses
```{r ,cache = TRUE}
data("ToothGrowth")

head(ToothGrowth)

str(ToothGrowth)

```

####Provide a basic summary of the data
Use the summary command to print a summary
```{r}
summary(ToothGrowth)
```
ToothGrowth is a dataframe with 60 observations on 3 variables. <br>

Let's chart the dose given and the length of tooth growth.

```{r,cache = TRUE, echo=FALSE, warning=FALSE, message=FALSE}
p <- ggplot(data = ToothGrowth, aes(x=dose,y=len )) + 
     geom_point() +facet_wrap(~supp) + geom_smooth() +
     labs(title ="ToothGrowth Factored by Supplement Type")
p
```

####Use confidence intervals and/or hypothesis tests to compare tooth growth

Perform a t-test to compare the length of the growth with the supplement.
```{r, cache=TRUE}
t.test(len ~ supp, paired = FALSE,  data = ToothGrowth)
```
Perform a t-test to compare the length of the growth with the dose given.
```{r}
t.test(ToothGrowth$len , ToothGrowth$dose)
```

Perform an analysis of the variance
```{r,cache = TRUE}
summary(aov(len ~ supp * dose, data=ToothGrowth))
```
####State your conclusions and the assumptions needed for your conclusions

The results for the analysis of length and the supplement show that the t-statistic is equal to 1.91 and the p-value is 0.06063. Since the p-value is very low, there is strong evidence of a mean increase in tooth length between the OJ and VC supplement.

The results for teh analysis of the length and the dose size show a large t statistic and a p value of almost zero.  This shows little significance between these two measures.
