---
title: "Statistical Inference Project - Coursera"
author: "Josh Starkey"
date: "January 31, 2016"
output: html_document
---

###Overview
This project you will investigate the exponential distribution in R and compare it with the Central Limit Theorem.   

First, let's load libraries
```{r, warning=FALSE, message=FALSE}
library(ggplot2)
```

###Part 1

####Simulations
Let's run the simulation of an exponential distribution with lambda = .2.

The expected mean of the distribution is 1/lambda.
```{r}
expmean <- 1 / .2
expmean
```
The expected variance of the distribution is also 1/lambda.
```{r}
expvar <- ((1 / .2)^2)
expvar
```
The number of simulations for the experiment is 1000
The number of samples is 40
```{r}
nosim <- 1000
nosample <- 40
```
Let's perform 1000 simulations of the exponential distribution.  We'll set the seed for reproducability.  
```{r, cache = TRUE, warning=FALSE,message=FALSE}
set.seed(1623);
meandata <- data.frame(
     x = c(apply(matrix(rexp(nosim * nosample, .2), nosim), 1, mean)),
     n = factor(rep(c(nosample), c(nosample))) 
     )
```

```{r, cache = TRUE, warning=FALSE,message=FALSE, echo=FALSE}
g <- ggplot(data = meandata, aes(x = x) ) +
     geom_histogram(aes(y = ..density..), fill = ("blue"), col=("darkgreen"), alpha=.4) +
     labs(title ="Average Means of 40 Samples", x="means") +
     geom_vline(xintercept=expmean)
g
```

Let's perform 1000 simulations of the exponential distribution.  We'll set the seed for reproducability.  
```{r, cache = TRUE, warning=FALSE,message=FALSE}
set.seed(1623);
vardata <- data.frame(
     x = c(apply(matrix(rexp(nosim * nosample, .2), nosim), 1, var)),
     n = factor(rep(c(nosample), c(nosample))) 
     )
```

```{r, cache = TRUE, warning=FALSE,message=FALSE, echo=FALSE}
g <- ggplot(data = vardata, aes(x = x) ) +
     geom_histogram(aes(y = ..density..), fill = ("blue"), col=("darkgreen"), alpha=.4) +
     labs(title ="Average Vars of 40 Samples", x="Vars") +
     geom_vline(xintercept=expvar)
g
```

####Sample Mean versus Theoretical Mean

I've added a vertical line at x = expmean = 5

####Sample Variance versus Theoretical Variance

I've added a vertical line at x = expvar = 25

####Distribution

###Part 2

####Load the ToothGrowth data and perform some basic exploratory data analyses
```{r ,cache = TRUE}
data("ToothGrowth")

head(ToothGrowth)

str(ToothGrowth)

```

####Provide a basic summary of the data
Use the summary command to print a summary
```{r}
summary(ToothGrowth)
```
The response is the length of odontoblasts (cells responsible for tooth growth) in 60 guinea pigs. Each animal received one of three dose levels of vitamin C (0.5, 1, and 2 mg/day) by one of two delivery methods, (orange juice or ascorbic acid (a form of vitamin C and coded as VC).

A data frame with 60 observations on 3 variables.

[,1]	len	numeric	Tooth length
[,2]	supp	factor	Supplement type (VC or OJ).
[,3]	dose	numeric	Dose in milligrams/day

```{r,cache = TRUE, echo=FALSE, warning=FALSE, message=FALSE}
p <- ggplot(data = ToothGrowth, aes(x=dose,y=len )) + 
     geom_point() +facet_wrap(~supp) + geom_smooth() 
p
```

####Use confidence intervals and/or hypothesis tests to compare tooth growth

Perform a paired t-test
```{r, cache=TRUE}
with(ToothGrowth, pairwise.t.test(len, dose))
```

Perform an analysis of the variance
```{r,cache = TRUE}
summary(aov(len ~ supp * dose, data=ToothGrowth))
```
####State your conclusions and the assumptions needed for your conclusions

Analysis here
