---
title: "Statistical Inference Project - Coursera"
author: "Josh Starkey"
date: "January 31, 2016"
output: html_document
---

##Overview
This project you will investigate the exponential distribution in R and compare it with the Central Limit Theorem.   

First, let's load libraries
```{r, warning=FALSE, message=FALSE}
library(ggplot2)
```

##Part 1

###Simulations
Let's run the simulation of an exponential distribution with lambda = .2.

The expected mean of the distribution is 1/lambda.
```{r}
expmean <- 1 / .2
expmean
```

Now let's perform 1000 simulations of the exponential distribution.  We'll set the seed for reproducability.  
```{r, cache = TRUE, warning=FALSE,message=FALSE}
set.seed(1623); mns = NULL; 
for (i in 1 : 1000) mns = c(mns, mean(rexp(40,.2))); 
```

```{r, cache = TRUE, warning=FALSE,message=FALSE, echo=FALSE}
p <- qplot(mns,
       geom="histogram",
       main = "Average Means of 40 Samples",
       xlab = "Means",
       fill=I("blue"),
       col=I("darkgreen"),
       alpha=I(.2)
 )
p + geom_vline(xintercept=5)
```

###Sample Mean versus Theoretical Mean

I've added a vertical line at x = expmean = 5

###Sample Variance versus Theoretical Variance

###Distribution

##Part 2

###Load the ToothGrowth data and perform some basic exploratory data analyses

###Provide a basic summary of the data

###Use confidence intervals and/or hypothesis tests to compare tooth growth

###State your conclusions and the assumptions needed for your conclusions

